---
title: "Final Project (League of Legends Match Dataset Analysis)"
author: "Guillermo"
date: "2026-05-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

```

```{r}
library(tidyverse)
library(janitor)
```

<!-- # 1. Import the data -->

```{r}
league <- read_csv("league_data.csv", show_col_types = FALSE)
league <- clean_names(league)
```

<!-- # 2. Data Cleaning -->

```{r}
league_clean <- league %>%
  filter(game_mode == "CLASSIC") %>%
  distinct() %>% 
  filter(!is.na(win)) %>%
  filter(individual_position != "Invalid") %>%
  select(-game_version) %>%
  group_by(puuid) %>%
  mutate(
    summoner_name = ifelse(
      is.na(summoner_name) | summoner_name == "",
      paste0("Unknown_", cur_group_id()),
      summoner_name
    )
  ) %>%
  ungroup() %>%
  mutate(
    win = as.logical(win),
    win_numeric = ifelse(win == TRUE, 1, 0),
    game_duration_min = game_duration / 60,
    kda = (kills + assists) / ifelse(deaths == 0, 1, deaths)
  )
```

# 3. Data Manipulation

### 3.a Summary Role

```{r}
role_summary <- league_clean %>%
  group_by(individual_position) %>%
  summarise(
    games = n(),
    win_rate = round(mean(win_numeric) * 100, 2),
    avg_kills = round(mean(kills), 2),
    avg_deaths = round(mean(deaths), 2),
    avg_assists = round(mean(assists), 2),
    avg_damage = round(mean(total_damage_dealt_to_champions), 2),
    avg_gold = round(mean(gold_earned), 2)
  ) %>%
  arrange(desc(win_rate))

role_summary
```

I analyzed the average performance metrics for each role, including win rate, kills, assists, damage, and gold earned. This helped identify which roles performed better overall in the dataset.

### 3.b Winrate per Champion and Role

```{r}
best_champions <- league_clean %>%
  group_by(individual_position, champion_name) %>%
  summarise(
    games = n(),
    win_rate = round(mean(win_numeric) * 100, 2),
    .groups = "drop"
  ) %>%
  filter(games >= 100) %>%
  group_by(individual_position) %>%
  slice_max(order_by = win_rate, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(desc(win_rate))

best_champions
```

I identified the champion with the highest win rate in each role using champions with at least 50 games played to avoid unreliable results from small sample sizes.

### 3.c Average KDA per role for winning players

```{r}
winner_kda_by_role <- league_clean %>%
  filter(win == TRUE) %>%
  group_by(individual_position) %>%
  summarise(
    avg_kda = round(mean(kda), 2),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_kda))

winner_kda_by_role
```

### 3.d Average KDA per role For Loser Players

```{r}
loser_kda_by_role <- league_clean %>%
  filter(win == FALSE) %>%
  group_by(individual_position) %>%
  summarise(
    avg_kda = round(mean(kda), 2),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_kda))

loser_kda_by_role
```

I compared the average KDA of winners and losers across each role to better understand how player performance relates to winning games.

### 3.e Percentages of Win and Losses

```{r}
win_loss_percentage <- league_clean %>%
  group_by(win) %>%
  summarise(
    total_games = n(),
    .groups = "drop"
  ) %>%
  mutate(
    percentage = round((total_games / sum(total_games)) * 100, 2)
  )

win_loss_percentage
```

I calculated the percentage of games won and lost to better understand the overall distribution of match outcomes in the dataset.

# 4. Statistical Analysis

### 4.a Average Game Duration

```{r}
avg_game_duration <- round(mean(league_clean$game_duration_min), 2)

avg_game_duration
```

### 4.b Average Games Played Per Day

```{r}
games_per_day <- league_clean %>%
  mutate(date = as.Date(game_start_utc)) %>%
  group_by(date) %>%
  summarise(
    total_games = n(),
    .groups = "drop"
  )

avg_games_per_day <- round(mean(games_per_day$total_games), 2)

avg_games_per_day
```

I calculated the average game duration and the average number of games played per day to better understand the general activity and match length in the dataset.

### 4.c Longest and Shortest Game

```{r}


longest_game <- round(max(league_clean$game_duration_min), 2)

shortest_game <- round(min(league_clean$game_duration_min), 2)

longest_game
shortest_game
```

I identified the longest and shortest matches in the dataset to better understand the range of game durations.

### 4.d Hypothesis Test

**Null Hypothesis:** Winning and losing players deal the same average damage.

**Alternative Hypothesis:** Winning and losing players deal different average damage.

```{r}

damage_test <- t.test(
  total_damage_dealt_to_champions ~ win,
  data = league_clean
)

damage_test
```

I used a t-test to compare the average champion damage between winners and losers. The p-value was less than 0.05, so the null hypothesis was rejected. This suggests that there is a statistically significant difference in damage dealt between winning and losing players. On average, winning players dealt more damage than losing players.

### 4.e Linear Regression

```{r}
gold_model <- lm(
  gold_earned ~ total_damage_dealt_to_champions,
  data = league_clean
)

summary(gold_model)
```

I used linear regression to analyze the relationship between champion damage and gold earned. The regression results showed a strong positive relationship between the two variables. The R-squared value was approximately 0.72, meaning that around 72% of the variation in gold earned can be explained by champion damage. The p-value was less than 0.05, indicating that the relationship was statistically significant.

# 5. Data Visualization

### 5.a Relationship Between Damage and Gold Earned

```{r}
gold_damage_chart <- league_clean %>%
  sample_n(3000) %>%
  ggplot(
    aes(
      x = total_damage_dealt_to_champions,
      y = gold_earned
    )
  ) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Relationship Between Damage and Gold Earned",
    subtitle = "Sample of 3,000 games",
    x = "Damage Dealt to Champions",
    y = "Gold Earned"
  ) +
  theme_minimal()

gold_damage_chart
```

This scatter plot visualizes the relationship between champion damage and gold earned. The graph shows that players who deal more damage generally earn more gold.

### 5.b Wins and Losses Percentage

```{r}
win_loss_chart <- ggplot(
  win_loss_percentage,
  aes(x = ifelse(win, "Win", "Loss"), y = percentage, fill = win)
) +
  geom_col(width = 0.5, color = "white") +
  geom_text(
    aes(label = paste0(percentage, "%")),
    vjust = -0.5,
    size = 4
  ) +
  scale_fill_manual(values = c("TRUE" = "#4CAF50", "FALSE" = "#F44336"), guide = "none") +
  labs(
    title = "Wins and Losses Percentage",
    subtitle = "By Team",
    x = "Team",
    y = "Win Percentage"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )

win_loss_chart
```

This visualization compares the percentage of wins and losses in the dataset.

### 5.c Role Distribution Percentage


```{r}
role_percentage <- league_clean %>%
  group_by(individual_position) %>%
  summarise(
    total_games = n(),
    .groups = "drop"
  ) %>%
  mutate(
    percentage = round((total_games / sum(total_games)) * 100, 2),
    label = paste0(percentage, "%")
  )

role_pie_chart <- ggplot(
  role_percentage,
  aes(x = "", y = percentage, fill = individual_position)
) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 4
  ) +
  labs(
    title = "Role Distribution Percentage",
    fill = "Role"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_void() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(face = "bold")
  )

role_pie_chart
```

This pie chart visualizes the percentage distribution of each role in the dataset.

### 5.d Damage Type Distribution

```{r}
damage_types <- tibble(
  damage_type = c("Physical", "Magic", "True"),
  total_damage = c(
    sum(league_clean$physical_damage_dealt_to_champions),
    sum(league_clean$magic_damage_dealt_to_champions),
    sum(league_clean$true_damage_dealt_to_champions)
  )
) %>%
  mutate(
    percentage = round((total_damage / sum(total_damage)) * 100, 2),
    label = paste0(percentage, "%")
  )

damage_type_chart <- ggplot(
  damage_types,
  aes(x = "", y = percentage, fill = damage_type)
) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 5
  ) +
  labs(
    title = "Damage Type Distribution",
    fill = "Damage Type"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_void() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(face = "bold")
  )

damage_type_chart
```

This chart visualizes the percentage distribution of physical, magic, and true damage in the dataset.
